---
title: "Effects of correlation levels on linear regression estimators: A simulation study."
author: "Sofía Villalobos Aliste"
date: "23-10-2021"
output: html_document
---

# 1. Aim

In this document, we will compare the mean square errors of the parameters of a linear regression model with two predictors drawn from a sample of n = 26, in two situations: First, when the two variables have a level of correlation of 0.1, and when they have a level of correlation of 0.9.


# 2. Set up  

### 2.1 Fixing random seed.  
Set the seed for replicable results.
```{r}
set.seed(161026768)
```

### 2.2 Number of simulations.  
Here we define the number of replications.
```{r}
reps<-1000 
```


### 2.3 Loading packages
```{r message=FALSE, warning=FALSE}
library(MASS)
library(kableExtra)
```


# 3. Parameters and storage
```{r}
b1<-5.8 #Coefficient beta 1
b2<-4   #Coefficient beta 2
n<-26   #Sample size
r1<-.1  #Correlation level
r9<-.9  #Correlation level

par.est1<-matrix(NA,nrow=reps,ncol=6) #Storage for correlation =.1
par.est9<-matrix(NA,nrow=reps,ncol=6) #Storage for correlation =.2
```



# 4. Draw samples 

### 4.1 Draw samples for correlation = 0.1
```{r}
data<-mvrnorm(n=n, mu=c(0, 0), Sigma=matrix(c(1, r1, r1, 1), nrow=2), empirical=TRUE)
V1<-data[, 1]  # standard normal (mu=0, sd=1)
V2<-data[, 2]

for(i in 1:reps){ 
  Y<-V1*b1+V2*b2+rnorm(n,0,1) #The true DGP, with N(0,1) error
  model1<-lm(Y~V1+V2) 
  vcv1<-vcov(model1)
  par.est1[i,1]<-model1$coef[1] 
  par.est1[i,2]<-model1$coef[2] 
  par.est1[i,3]<-model1$coef[3]
  
  par.est1[i,4]<-sqrt(diag(vcv1)[1]) 
  par.est1[i,5]<-sqrt(diag(vcv1)[2])
  par.est1[i,6]<-sqrt(diag(vcv1)[3])
}
```

### 4.2 Draw sample for correlation = 0.9
```{r}
data<-mvrnorm(n=n, mu=c(0, 0), Sigma=matrix(c(1, r9, r9, 1), nrow=2), empirical=TRUE)
V1<-data[, 1]  # standard normal (mu=0, sd=1)
V2<-data[, 2]

for(i in 1:reps){ 
  Y<-V1*b1+V2*b2+rnorm(n,0,1) 
  model9<-lm(Y~V1+V2) 
  vcv9<-vcov(model9)
  par.est9[i,1]<-model9$coef[1] #INTERCEPTO
  par.est9[i,2]<-model9$coef[2] 
  par.est9[i,3]<-model9$coef[3]
  
  par.est9[i,4]<-sqrt(diag(vcv9)[1]) #SE INTERCEPTO
  par.est9[i,5]<-sqrt(diag(vcv9)[2])
  par.est9[i,6]<-sqrt(diag(vcv9)[3])
}
```


# 5. Assessment: Mean square error 
```{r}
#MSE for Correlation = 0.1
mse.beta1_c1<-mean((par.est1[,2]-b1)^2)
mse.beta2_c1<-mean((par.est1[,3]-b2)^2)

#MSE for correlation = 0.9
mse.beta1_c9<-mean((par.est9[,2]-b1)^2)
mse.beta2_c9<-mean((par.est9[,3]-b2)^2)

corr01<-cbind("0.1",mse.beta1_c1,mse.beta2_c1)
corr09<-cbind("0.9",mse.beta1_c9,mse.beta2_c9)

output<-rbind(corr01,corr09)
colnames(output)<-c("Corr","B1","B2")
kable(output,"html",digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),full_width = T)
```

# 6. Replication

Set a different seed for replication.
```{r}
set.seed(834844)
```

```{r}
data<-mvrnorm(n=n, mu=c(0, 0), Sigma=matrix(c(1, r1, r1, 1), nrow=2), empirical=TRUE)
V1<-data[, 1]  # standard normal (mu=0, sd=1)
V2<-data[, 2]

for(i in 1:reps){ 
  Y<-V1*b1+V2*b2+rnorm(n,0,1) #The true DGP, with N(0,1) error
  model1<-lm(Y~V1+V2) 
  vcv1<-vcov(model1)
  par.est1[i,1]<-model1$coef[1] 
  par.est1[i,2]<-model1$coef[2] 
  par.est1[i,3]<-model1$coef[3]
  
  par.est1[i,4]<-sqrt(diag(vcv1)[1]) 
  par.est1[i,5]<-sqrt(diag(vcv1)[2])
  par.est1[i,6]<-sqrt(diag(vcv1)[3])
}
data<-mvrnorm(n=n, mu=c(0, 0), Sigma=matrix(c(1, r9, r9, 1), nrow=2), empirical=TRUE)
V1<-data[, 1]  # standard normal (mu=0, sd=1)
V2<-data[, 2]

for(i in 1:reps){ 
  Y<-V1*b1+V2*b2+rnorm(n,0,1) 
  model9<-lm(Y~V1+V2) 
  vcv9<-vcov(model9)
  par.est9[i,1]<-model9$coef[1] #INTERCEPTO
  par.est9[i,2]<-model9$coef[2] 
  par.est9[i,3]<-model9$coef[3]
  
  par.est9[i,4]<-sqrt(diag(vcv9)[1]) #SE INTERCEPTO
  par.est9[i,5]<-sqrt(diag(vcv9)[2])
  par.est9[i,6]<-sqrt(diag(vcv9)[3])
}
mse.beta1_c1<-mean((par.est1[,2]-b1)^2)
mse.beta2_c1<-mean((par.est1[,3]-b2)^2)

#MSE for correlation = 0.9
mse.beta1_c9<-mean((par.est9[,2]-b1)^2)
mse.beta2_c9<-mean((par.est9[,3]-b2)^2)

corr01<-cbind("0.1",mse.beta1_c1,mse.beta2_c1)
corr09<-cbind("0.9",mse.beta1_c9,mse.beta2_c9)

output<-rbind(corr01,corr09)
colnames(output)<-c("Corr","B1","B2")
kable(output,"html",digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),full_width = T)
```

We can see that the results are similar to the ones obtained in the original study.

# END OF DOCUMENT


```{r}
sessionInfo()
```

